Learning CNN filters from user-drawn image makers for coconut-tree image classification

### 2020-06-12 main.tex
review definitions

The user selects a few training images to compose a very small dataset $\D$ and draws labeled markers in image regions that best discriminate the classes. The convolution between an image $I$ and a filter $F\in {\cal F}$ generates positive activations in regions whose local patterns are detected by $F$. We wish then to estimate the weights of $F$ such that those local patterns are characteristic of one given class. For sake of clarity, one must interpret the convolution operation at a pixel $p$ of $I$ as the inner product $\vect(P_I(p)) \cdot  \vect(F)$, where $P_I(p)$ is a patch with $k\times k$ pixels and $m$ image channels around $p$, $F$ is a filter with $k\times k\times m$ weights, and $\vect$ is the vectorization operation. A filter $F$ can discriminate a class $i\in \{1,2,\ldots,c\}$ among $c$ classes when it generates positive results for local patterns $\vect(P_I(p))$ of class $i$ and negative results for patterns from other classes. That is, $\vect(F)$ is the normal vector of a hyperplane in $\Re^{k\times k\times m}$ which detects the patterns of class $i$ at its positive side. We wish to estimate ${\cal F}$ such that its filters will detect patterns from all classes in different positions $p$.

For a problem with $c$ classes where $\lambda(p)=i\in \{1,2,\ldots,c\}$ is the label of a marker pixel $p$ from class $i$, let ${\cal M}_I$ be a set of marker pixels drawn in image $I\in {\cal D}$ and $P_I(p)$ be the respective patch around a pixel $p\in {\cal M}_I$. Let ${\cal P}$ be the set of all patches around marker pixels from all images $I\in {\cal D}$, with representative examples  $\vect(P_I(p))$ from all the $c$ classes.  
\begin{eqnarray}
{\cal P} & = & \bigcup_{I\in {\cal D},p\in {\cal M}_I, \lambda(p)\in [1,c]} P_I(p)
\label{eq.pset}
\end{eqnarray}
The patch set ${\cal P}$ defines samples from all classes in $\Re^{k\times k\times m}$ and a clustering operation on ${\cal P}$ creates groups with similar local patterns. Ideally, each group should contain patterns of a same class. However, some classes might share similar patterns. We should at least expect that their patterns be detected in different image regions. The groups must be shifted to the origin of $\Re^{k\times k\times m}$ by subtracting the mean value of each feature in ${\cal P}$. Additionally, the standard deviation of each feature is computed and used for standardization. This operation allows batch normalization of image sets using the mean and standard deviation of the features from the image markers -- i.e.,  \emph{a marker-based batch normalization}. The centroid of each cluster defines the weights of each filter $F\in {\cal F}$ and we force norm $\|\vect(F)\|=1$ to avoid preferences among them. The centralization for filter definition is paramount to eliminate activations in regions from other classes, whose patterns fall in the negative side of the hyperplane with normal $\vect(F)$. Figure~\ref{fig:filter}a illustrates three groups with color-coded samples from two classes in a hypothetical 2D feature space. The marker-based batch normalization and convolution with filters obtained as centroids of the three groups creates a new 3D feature space (Figure~\ref{fig:filter}b), in which the classes can be more easily detected by two filters.

>>>>

The user selects a few training images to compose a very small dataset $\D$ and draws labeled markers in image regions that best discriminate the classes. The convolution between an image $I$ and a filter $F\in {\cal F}$ generates positive activations in regions whose local patterns are detected by $F$. We wish then to estimate the weights of $F$ such that those local patterns are characteristic of one given class. For sake of clarity, one must interpret the convolution operation at a pixel $p$ of $I$ as the inner product $\vect(P_I(p)) \cdot \vect(F)$, where $P_I(p)$ is a $k\times k$ patch with $m$ channels around $p$, $F$ is a filter with shape $k\times k \times m$, and $\vect$ is the vectorization operation. A filter $F$ can discriminate a class $i \in \{1,2,\ldots,c\}$ among $c$ classes when it generates positive results for local patterns $\vect(P_I(p))$ of class $i$ and negative results for patterns from other classes. That is, $\vect(F)$ is the normal vector of a hyperplane in $\R^{k\times k\times m}$ which detects the patterns of class $i$ at its positive side. We wish to estimate ${\cal F}$ such that its filters will detect patterns from all classes in different positions $p$.

\textcolor{blue}{For a problem with $c$ classes where $\lambda(p)=i\in \{1,2,\ldots,c\}$ is the label of a marker pixel $p$ from class $i$, let ${\cal M}_I$ be a set of marker pixels drawn in image $I\in {\cal D}$ and $P_I(p)$ be the respective patch around a pixel $p\in {\cal M}_I$. Let ${\cal P}$ be the set of all patches around marker pixels from all images $I\in {\cal D}$, with representative examples  $\vect(P_I(p))$ from all the $c$ classes. We define the set of patches from class $i$ as
\begin{eqnarray}
{\cal P}_i & = & \bigcup_{I\in {\cal D},p\in {\cal M}_I, \lambda(p) = i}{P_I(p)}.
\label{eq:pset}
\end{eqnarray}
A clustering operation on ${\cal P}$ creates groups with similar local patterns. Ideally, each group should contain patterns of class $c$. However, some classes might share similar patterns. We should at least expect that their patterns be detected in different image regions. The groups must be shifted to the origin of $\R^{k\times k\times m}$ by subtracting the mean value of each patch in ${\cal P}$, where $\mathcal{P} = \bigcup_{i \in \{1, 2, \ldots, c\}}{\mathcal{P}_i}$. Additionally, the standard deviation of each patch is computed and used for standardization. This operation allows batch normalization of image sets using the mean and standard deviation of the patches from the image markers -- i.e.,  \emph{a marker-based batch normalization}. The centroid of each cluster defines the weights of each filter $F\in {\cal F}$ and we force norm $\|\vect(F)\|=1$ to avoid preferences among them. The centralization for filter definition is paramount to eliminate activations in regions from other classes, whose patterns fall in the negative side of the hyperplane with normal $\vect(F)$. Figure~\ref{fig:filter} illustrates three groups with color-coded samples from two classes in a hypothetical 2D feature space. The marker-based batch normalization and convolution with filters obtained as centroids of the three groups creates a new 3D feature space (Figure~\ref{fig:filter}b), in which the classes can be more easily detected by two filters.}
